### 2017-09-14 doc/msi-api.tex
adding msi_matrix_getField

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{API Overview}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The MSI API's are provided in the file ``\texttt{msi.h}". Throughout this section, unless specified, DOF, matrix row and column ID's are specified by a local ID. A word ``\emph{global}" in a function name indicates that the function involves global operation or global data.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Naming Convention}

MSI API function name consists of three words connected with `$\_$'.

\begin{itemize}
\item the first word is ``msi"
\item the second word is an operation target. If the operation target is system-wide, the operation target is ommited. For instance, the function name which initializes the MSI service consists of two words: \emph{msi}$\_$\emph{start}.
\item the third word is the operation description starting with a verb. For example, the function \emph{msi}$\_$\emph{matrix}$\_$\emph{getNumIter} returns the number of iterations in the solve operation. 
\end{itemize}

The following are operation targets used in the second word.

\begin{itemize}
\item \textit{field}: the api is performed on a field object
\item \textit{matrix}: the api is performed on a Petsc matrix
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Abbreviation}

Abbreviations may be used in API naming. See \textit{http://scorec.rpi.edu/wiki/Abbriviations} for more information.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Data Types and Classes}

For a geometry, partition and mesh model, the term \emph{instance} is used to indicate the model data existing on each process. For example, a mesh instance on process \emph{i} means a pointer to a mesh data structure on process \emph{i}, in which all mesh entities on process \emph{i} are contained and from which they are accessible. For all other data such as field and matrix, the term \emph{handle} is used to indicate the pointer to the data. For example, a matrix handle means a pointer to the matrix. The predefined data type has a prefix \emph{p} to indicate the pointer data type.

The following are predefined data types used in the interface function parameters.

\begin{tabular}{lp{14cm}}	
	pMesh  & 	a mesh instance (defined in ``\texttt{pumi.h}")\\
        pMeshEnt & 	a mesh entity handle (defined in ``\texttt{pumi.h}")\\
        pField & 	a field handle (defined in ``\texttt{pumi.h}")\\
	pOwnership & 	a user-defined ownership handle (defined in ``\texttt{pumi.h}") \\
 	pMatrix &	a matrix handle \\
\end{tabular}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Enumeration Types}

The enumeration type for matrix type is:

\begin{verbatim}
    msi_matrix_type {
        MSI_MULTIPLY = 0 /* matrix for multiplication */,
        MSI_SOLVE      	 /* 1 - matrix for solve */
    }
\end{verbatim}\vspace{-1cm}\hspace{1cm}

The enumeration type for matrix status is:

\begin{verbatim}
    msi_matrix_status {
        MSI_NOT_FIXED = 0 /* matrix is modifiable */,
        MSI_FIXED         /* 1 - matrix is not modifiable */
    }
\end{verbatim}\vspace{-1cm}\hspace{1cm}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{API Functions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Initialization and Finalization}

The functions initialize/finalize the MSI operations.

\begin{verbatim}
void msi_start(
    pMesh  /* in */  m, 
    pOwnership  /* in */  o=NULL, 
    pShape  /* in */  s=NULL)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a mesh, an ownership handle, and a field shape handle, initialize MSI services for solver-PUMI interactions. The ownership and field shape are optional (default: $NULL$). Ownership rule determines the owner copy among duplicated mesh entity copies by partitioning or ghosting. If the ownership handle is $NULL$, the PUMI's default rule is used (a part with the minimum process rank is the owning part of duplicate copies). Field shape defines the node distribution where the coordinates and field values (DOF's) are stored. If the shape is \emph{NULL}, the field shape saved in mesh (returned by \texttt{pumi$\_$mesh$\_$getShape}) is used. If it is not \textit{NULL}, the field shape of the mesh is set to $s$.

Combining ownership rule and field shape, local/global node numbering are generated.
See PUMI User's Guide for more information on ownership rule, field shape and numberings.

Note that the following operations should be performed prior to this function.
\begin{itemize}
\item MPI initialization
\item Solver initialization (e.g. PetscInitialize)
\item PUMI initialization 
\item model and mesh loading
\end{itemize}

\begin{verbatim}
int msi_finalize()
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Finalize the MSI services and clears all internal data. Note that the following operations should follow to complete further finalizations.
\begin{itemize}
\item mesh deletion
\item PUMI finalization
\item Solver finalization (e.g. PetscFinalize)
\item MPI finalization
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Field}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The attributes of PUMI field are:
\begin{itemize}
\item field name: unique string 
\item field size: the number of field data per node
\item field type: \texttt{PUMI$\_$SCALAR} for a real number, \texttt{PUMI$\_$VECTOR} for real numbers of size 3, \texttt{PUMI$\_$MATRIX} for real numbers of size 9, and \texttt{PUMI$\_$PACKED} for real numbers of arbitrary size
\item shape function to specify node distribution. The default shape is \textit{Linear} where nodes are associated only with vertices.
\end{itemize}

To support solver with advanced features, MSI supports the field API's with multiple field variables per node and complex number. If not specified here, use the field API's \texttt{PUMI.h} for the rest of field operations.

\begin{verbatim}
pField msi_field_create (
    pMesh  /* in */  m,
    const char*  /* in */  field_name, 
    int  /* in */  nv, 
    int  /* in */  nd, 
    pShape  /* in */  s=NULL)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a mesh instance, field name, the number of values per node ($nv$), the number of field data per value of node ($nd$), and a shape function, \emph{(i)} create a field for all nodes (owned, non-owned part boundary and ghost), \emph{(ii)} switch the memory space for field data to a contiguous array, and \emph{(iii)} initialize the field data. The \emph{size}, the number of field data per node is \textit{nv}$*$\textit{nd}. The size of contiguous array for each process is \textit{size}$*$\textit{nn} for real number and \textit{size}$*$\textit{nn}$*$2 for complex number, where \textit{nn} is the number of local nodes on each process. The field data type (real or complex) is determined at time of configuration. See Section~\ref{install} for how to configure MSI with complex number.

The field shape is optional (default: \textit{NULL}). If it is \textit{NULL}, the field shape of the mesh is used (a field shape returned by \texttt{pumi$\_$mesh$\_$getShape}).

 If $nv$ is 1, \texttt{msi$\_$field$\_$create (m, field$\_$name, nv, nd, s)} is equivalent to \texttt{pumi$\_$field$\_$create (m, field$\_$name, nd, PUMI$\_$PACKED, s)} followed by \texttt{pumi$\_$field$\_$freeze} and \texttt{pumi$\_$ment$\_$setField} with value $0$ for all nodes.

\begin{verbatim}
int msi_field_getNumVal (pField  /* in*/  f)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, return the number of value per node, \textit{nv}. If the field was created with \texttt{pumi$\_$field$\_$create}, return 1.

\begin{verbatim}
int msi_field_getSize (pField  /* in*/  f)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, return the field size, \textit{nv}$*$\textit{nd}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Node}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{verbatim} 
int msi_node_getID (
    pMeshEnt  /* in */  e, 
    int  /* in */  n)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a mesh entity handle and the index of node $n$, return local number of $n^{th}$ node of the mesh entity.

\begin{verbatim}
int msi_node_getGlobalID (
    pMeshEnt  /* in */  e, 
    int  /* in */  n)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a mesh entity handle and the index of node $n$, return global number of $n^{th}$ node of the mesh entity.
The global numbering is based on the ownership handle provided in \texttt{msi$\_$start}.

\begin{verbatim}
void msi_node_setField (
    pField  /* in */  f, 
    pMeshEnt  /* in */  e, 
    int  /* in */  n,
    int  /* in */  size_dof, 
    double*  /* in */  dof_data
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, a mesh entity handle, the index of node $n$, size of DOF data and a double array containing DOF data, set the field data (DOF) of $n^{th}$ node of the entity.

\begin{verbatim}
int msi_node_getField(
    pField  /* in */  f, 
    pMeshEnt  /* in */  e, 
    int  /* in */  n,
    double*  /* out */  dof_data)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, a mesh entity handle, the index of node $n$, fill the array \textit{dof$\_$data} with field data (DOF) of $n^{th}$ node of the entity.

\begin{verbatim}
void msi_node_getFieldID (
    pField  /* in */  f, 
    pMeshEnt  /* in */  e, 
    int  /* in */  n,
    int*  /* out */  start_dof_id, 
    int*  /* out */  end_dof_id_plus_one)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, a mesh entity handle, and the index of node $n$, return the starting local ID and the ending local ID plus one for field data (DOF) of $n^{th}$ node of the entity.

\begin{verbatim}
void msi_ment_getGlobalFieldID (
    pField  /* in */  f, 
    pMeshEnt  /* in */  e, 
    int  /* in */  n,
    int*  /* out */  start_dof_id, 
    int*  /* out */  end_dof_id_plus_one)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a field handle, a mesh entity handle, and the index of node $n$, return the starting global ID and the ending global ID plus one for field data (DOF) of $n^{th}$ node of the mesh entity.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{PETSc Matrix and Solver}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{verbatim}
pMatrix msi_matrix_create (
        int  /* in */  matrix_type,
        pField  /* in */  field)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given matrix type (0 for matrix-vector multiplication, 1 for solver) and a field handle, create a matrix and return its handle. The input field handle is used to set up the matrix size  and row/column ID's.The initial status of matrix is \textit{MSI$\_$NOT$\_$FIXED} so the matrix values can be modified.

\begin{verbatim}
void msi_matrix_delete (pMatrix  /* in */  matrix)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, delete the matrix.

\begin{verbatim}
pField msi_matrix_getField (pMatrix  /* in */  matrix)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, return the field handle associated with the matrix.

\begin{verbatim}
void msi_matrix_assemble (pMatrix  /* in */  matrix)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, perform matrix assembly and set the status of matrix  to \textit{MSI$\_$FIXED}. 
The matrix values cannot be modified any further.

\begin{verbatim}
void msi_matrix_insert (
    pMatrix  /* in */  matrix, 
    int  /* in */  r, 
    int  /* in */  c, 
    int  /* in */  scalar_type,
    double*  /* in */  value)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, local DOF ID $r$ and $c$, scalar type (0 for real, 1 for complex) and an array of double values, insert or overwrite \textit{value} to the matrix at (\textit{r},\textit{c}). If \textit{value} is a real number, \textit{scalar$\_$type} is 0. Otherwise, \textit{scalar$\_$type} is 1. A real type value can be inserted into a complex matrix but a complex type value cannot be inserted into a real matrix.

If the number of DOF's associated with the field is $d$, $r$ or $c$ divided by $d$ is a local ID of corresponding mesh entity. Use \texttt{msi$\_$ment$\_$getFieldID} to get local DOF ID of a mesh entity.

\begin{verbatim}
void msi_matrix_add (
    pMatrix  /* in */  matrix, 
    int  /* in */  r, 
    int  /* in */  c, 
    int  /* in */  scalar_type,
    double*  /* in */  value)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, local DOF ID $r$ and $c$, scalar type (0 for real, 1 for complex) and an array of double values, add \textit{value} to the existing value of matrix at (\textit{r},\textit{c}). If \textit{value} is a real number, \textit{scalar$\_$type} is 0. Otherwise, \textit{scalar$\_$type} is 1. A real type value can be added into a complex matrix but a complex type value cannot be added into a real matrix.

If the number of DOF's associated with the field is $d$, $r$ or $c$ divided by $d$ is a local ID of corresponding mesh entity. Use \texttt{msi$\_$ment$\_$getFieldID} to get local DOF ID of a mesh entity.

\begin{verbatim}
void msi_matrix_addBlock (
    pMatrix  /* in */  matrix, 
    int  /* in */  elem_id, 
    int  /* in */  r_value_index, 
    int  /* in */  c_value_index, 
    double*  /* in */  values)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, a local element ID, row value index, column value index, and a double array containing element matrix values, add values in the element matrix to the global matrix. The the number of the values associated with the field is 1, accepted value for \textit{r$\_$value$\_$index} and \textit{c$\_$value$\_$index} is 0.

\begin{verbatim}
void msi_matrix_setBC (
    pMatrix  /* in */  matrix, 
    int  /* in */  local_row_index)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle of type 1 (\textit{MSI$\_$SOLVE}) and a local row index, zero out all off-diagonal values in the row of the matrix and set the diagonal value to one. The operation is carried out during finalizing the matrix. It will overwrite other insertion operations to the local row of the matrix. For complex-valued matrix, the real part of the diagonal is set to one and the imaginary part is set to zero.
This function should be called on all processes that use the DOF numbering associated with the matrix row.

\begin{verbatim}
void msi_matrix_setLaplaceBC (
    pMatrix  /* in */  matrix, 
    int  /* in */  row, 
    int  /* in */  size, 
    int*  /* in */  columns, 
    double*  /* in */  values)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle of type 1 (\textit{MSI$\_$SOLVE}), a local row index, the  number of values to be inserted (\textit{size}), the columns to set the values (\textit{columns}), and the values to be set in the order of the \textit{columns}, set multiple values for the row of the matrix. If real values are inserted into a complex matrix, the corresponding imaginary parts are set to zero. The operation is carried out during finalizing the matrix. 
This function will overwrite other insertion operations to the row. This function should be called on all processes that use the DOF numbering associated with the matrix row.

\begin{verbatim}
void msi_matrix_multiply (
    pMatrix  /* in */  A, 
    pField  /* in */   x, 
    pField  /* out */  b)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle ($A$) of type 0 (\textit{MSI$\_$MULTIPLY}) and an input field handle ($x$), perform the matrix-vector multiplication \textit{``Ax}" and write the result in the field $b$. If the input matrix or the input field is complex-valued, the output field must be complex-valued.

\begin{verbatim}
void msi_matrix_solve (
        pMatrix  /* in */   A, 
        pField   /* in */   b, 
        pField   /* out */  x)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle ($A$) of type 1 (\textit{MSI$\_$SOLVE}) and a RHS field handle ($b$), solve the global discrete equation \textit{``Ax=b}" and write the solution into the field $x$.

\begin{verbatim}
int msi_matrix_getNumIter (pMatrix  /* in */  matrix)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle of type 1 (\textit{MSI$\_$SOLVE}), return the number of iterations of solve operation.

\begin{verbatim}
void msi_matrix_print(pMatrix  /* in */  matrix)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, print the $non$-$zero$ matrix value along with global row/colume index. The row/column ID starts with 0. PUMI provides an equivalent API for fields and nodes.

\begin{verbatim}
void msi_matrix_write (
    pMatrix  /* in */  matrix, 
    const char*  /* in */  file_name, 
    int  /* in */  start_index=0)
\end{verbatim}\vspace{-.5cm}\hspace{1cm}
Given a matrix handle, file name and a starting local ID for nodes (default is 0), write the $non$-$zero$ matrix values in file(s). For each process \textit{i}, the matrix information is written in ``\texttt{filename-i}". If \textit{file$\_$name} is \textit{NULL}, \texttt{msi$\_$matrix$\_$print} is performed.
PUMI provides an equivalent API for fields and nodes.